---
title: "Vis I"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


Making our first plot.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```


```{r}
ggp_weather_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
weather_df |> 
  filter(is.na(tmax))
```

## Fancier scatterplots!

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +  #applies color=name to everything that comes after this line of code
  geom_point(alpha = .3, size = .8) + ##alpha .3 makes it more transparent
  geom_smooth(se = FALSE)  #places smooth curve fit through the data points, se = false removes standard error bars
```

Where you define aesthetics can matter

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3, size = .8) + #color=name is placed differently here than the code chunk above, now only there is one smooth curve. 
  geom_smooth(se = FALSE)  
```

use faceting real quick - making multiple of the same plot varied by a variable

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) #left of ~ "." means everything; the right of the ~ refers to separate plots by what column; to better see what this line of code does, can run the code chunk with and without it and see the difference in plots
```

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

Learning assessment:
Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in `geom_smooth()`).

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr = tmax * (9/5) + 32,
    tmin_fahr = tmin * (9/5) + 32
  ) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) #straight line instead of curvy line
```

## Small things
```{r}
#plot the smooth curve without the datapoints
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = "blue")) + #this will output red color but call it blue because ggplot does not know blue and will make something up
  geom_point()
```

To actually have a blue plot
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "blue") 
```

## Univariate plots

```{r}
weather_df |> 
  ggplot(aes(x = tmin)) + 
  geom_histogram()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram(position = "dodge") #position = "dodge" puts bars next to each other rather than on top of each other
```

How would i fix this? maybe facet?

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

maybe a density plot?

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3) 
```

maybe a box plot?

```{r}
weather_df |> 
  ggplot(aes(x = name, y= tmin, fill = name)) +
  geom_boxplot()
```

violin plots

```{r}
#violin plots takes density plot and butterflies them, can show things that are skewed, bimodal, or multimodal well in terms of distribution
weather_df |> 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()
```

ridge plot

```{r}
#this plot might be useful when you have multiple things you want to compare at the same time
weather_df |> 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

Learning assessment: Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.
```{r}
weather_df |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

```{r}
weather_df |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .3) 

weather_df |> 
  ggplot(aes(x = name, y= prcp, fill = name)) +
  geom_boxplot()

weather_df |>
  filter(prcp > 10) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .3) 
  
```

```{r}
weather_df |> 
  ggplot(aes(x = name, y = prcp, fill = name)) +
  geom_violin()
```

```{r}
weather_df |> 
  ggplot(aes(x = prcp, y = name)) +
  geom_density_ridges()
```

## Saving and embedding plots

saving plots

```{r}
ggp_weather = 
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 6)
```

embedding plots

```{r, fig.width = 6, fig.asp = .6} 
#R decides the size of the embedded figure automatically, so can set the figure size
#fig.asp refers to aspect ratio. 
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
```

