---
title: "Vis II"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```



Make a scatterplot but fancy this time

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations."
  )
```

Scales -- start with `x` and `y` and then do `color`/

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations."
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 20),
    labels = c("-15", "0", "20")
  ) +
  scale_y_continuous(
    limits = c(0, 30),
    transform = "sqrt"
    )
```


Look at color:

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations."
  ) +
  viridis::scale_color_viridis(discrete = TRUE) #making viridis variable discrete not continuous
```



## Themes

```{r}
ggp_scatterplot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations."
  ) +
  viridis::scale_color_viridis(discrete = TRUE)
```

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom")
```

```{r}
ggp_scatterplot +
  theme_bw() +
  theme(legend.position = "bottom")
```

Order matters ...

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  theme_bw()
```


```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  theme_minimal() #same as theme_bw without the box around outside
```

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  theme_classic() #this looks to be what I like
```


```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  ggthemes::theme_excel() #does not look good
```

Learning Assessment: Revisit the plot showing tmax against date for each location. Use labels, scale options, and theme changes to improve the readability of this plot

New Scatterplot

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  viridis::scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Date",
    y = "Maximum Temp (C)",
    title = "Seasonal variation in Max Temp"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```



Extra bonus stuff in `ggplot`

Use different datasets in different `geom`s

```{r}
central_park_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df |> 
  filter(name == "Molokai_HI")

molokai_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() + #scatterplot for molokai 
  geom_line(data = central_park_df) #line plot for central park
```



## Multiple panels

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density() +
  facet_grid(. ~ name)
```


```{r}
ggp_tmax_tmin = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)

ggp_tmax_density = 
  weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)

ggp_tmax_date =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point () +
  geom_smooth(se = FALSE)

(ggp_tmax_tmin + ggp_tmax_density) / ggp_tmax_date #using patchwork
```




